---
title: "Project 6"
author: "Your Name Here"
format: 
  html:
    self-contained: true
    echo: true
    message: false
    warning: false
    error: true # IMPORTANT NOTE - this makes the file render even if there is an error. it is imperative that you check your .html document *before submission* to make sure that it has all results in it.
editor: source
---


**This week you will be analyzing data from the Jackson Heart Study (JHS). You can find the data in the Week 1 module on Canvas. For full credit, you must include all code chunks and R output backing up your responses. Remember to set your seed in all problems for reproducibility purposes.**

**1. Use leave-one-out cross validation to determine which model fits better:**

**1a. Systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), diabetic status (*diab3cat*; 0=non-diabetic, 1=pre-diabetic, 2=diabetic), body mass index (*bmi*, kg/m2), the interaction between diabetic status and age, and the interaction between diabetic status and education.**

```{r}

```

**1b. Systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), diabetic status (*diab3cat*; 0=non-diabetic, 1=pre-diabetic, 2=diabetic), body mass index (*bmi*, kg/m2).**

```{r}

```

**2. Consider the following model: diabetic status (*Diabetes*; 0=non-diabetic, 1=diabetic) as a function of age (*age*; years), weight (*weight*; kg), hypertension status (*HTN*; 0=normotensive, 1=hypertensive), health status as indicated by high density lipoproteins (*hdl3cat*; 0=poor health, 1=intermediate health, 2=ideal health), the interaction between weight and hypertension status, the interaction between weight and age, and the interaction between weight and health status as indicated by high density lipoproteins.**

```{r}

```

<!-- You want to time how long each CV below takes. hint: https://rdrr.io/cran/tictoc/man/tic.html -->

**2a. Perform 2-fold cross validation.**

```{r}

```

**2b. Perform 5-fold cross validation.**

```{r}

```

**2c. Perform 10-fold cross validation.**

```{r}

```

**2d. Perform 25-fold cross validation.**

```{r}

```

**2e. Perform 50-fold cross validation.**

```{r}

```

**2f. Perform 100-fold cross validation.**

```{r}

```

**2g. What did you observe, if anything, about the CV values?**

**2h. What did you observe, if anything, about the processing time?**

**3. Use 25-fold cross validation to determine which model fits better:**

**3a. Modeling the number of ideal health indicators (use blood pressure (*idealHealthBP*; 1=ideal health, 0=not ideal health), smoking status (*idealHealthSMK*; 1=ideal health, 0=not ideal health), diabetes (*idealHealthDM*; 1=ideal health, 0=not ideal health), diet  (idealHealthNutrition; 1=ideal health, 0=not ideal health), physical activity (*idealHealthPA*; 1=ideal health, 0=not ideal health), obesity  (*idealHealthBMI*; 1=ideal health, 0=not ideal health), and high cholesterol  (*idealHealthChol*; 1=ideal health, 0=not ideal health)) as a function of age (*age*; years), health insurance (*PrivatePublicIns*; 0=uninsured, 1=private insurance only, 2=public insurance only, 3=private and public insurances), education status (*HSgrad*; 0=did not graduate high school, 1=graduated high school), and the interaction between age and health insurance.**

```{r}

```

**3b. Modeling the number of ideal health indicators (use blood pressure (*idealHealthBP*; 1=ideal health, 0=not ideal health), smoking status (*idealHealthSMK*; 1=ideal health, 0=not ideal health), diabetes (*idealHealthDM*; 1=ideal health, 0=not ideal health), diet  (idealHealthNutrition; 1=ideal health, 0=not ideal health), physical activity (*idealHealthPA*; 1=ideal health, 0=not ideal health), obesity  (*idealHealthBMI*; 1=ideal health, 0=not ideal health), and high cholesterol  (*idealHealthChol*; 1=ideal health, 0=not ideal health)) as a function of age (*age*; years), health insurance (*PrivatePublicIns*; 0=uninsured, 1=private insurance only, 2=public insurance only, 3=private and public insurances), education status (*HSgrad*; 0=did not graduate high school, 1=graduated high school).**

```{r}

```

**4. Required for graduate students / extra credit for undergraduate students: this is a challenge question! If you cannot figure it out, please document the research you performed when searching for the answer and what you tried in terms of R code. Use leave-one-out cross validation to determine which model is better:**

**4a. Model diabetic status (*diab3cat*; 0=non-diabetic, 1=pre-diabetic, 2=diabetic) as a function of age (*age*; years), weight (*weight*; kg), hypertension (*HTN*; 1=yes, 0=no), high density lipoprotein (*hdl3cat*; 0=poor health, 1=intermediate health, 2=ideal health), the interaction between high density lipoprotein and age, and the interaction between weight and age.**

```{r}

```

**4b. Model diabetic status (*diab3cat*; 0=non-diabetic, 1=pre-diabetic, 2=diabetic) as a function of age (*age*; years), weight (*weight*; kg), hypertension (*HTN*; 1=yes, 0=no), and high density lipoprotein (*hdl3cat*; 0=poor health, 1=intermediate health, 2=ideal health).**

```{r}

```

**5. Required for graduate students / extra credit for undergraduate students: Pick the results from either Q1 or Q3 to write a summary paragraph for. This paragraph should outline the methodology + results and be readable for your supervisor (not a statistician or data scientist).**